Using Anchor Points to Define and Transfer Spatial Regions Based on Context

### 2012-04-13 code/scenarios/cdsr/trunk/papers/QR2012/main-QR12-cdsr.tex

\title{Towards a Cognitive System That Can Recognize Spatial Regions Based on Context}

\author{%
Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk 
\And % 
Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
\And % 
Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  
\AND % 
Graham S. Horn \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ gsh148@cs.bham.ac.uk 
\And % 
John D. Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a cognitive system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

Consider a janitorial robot cleaning a classroom. While performing this task, it encounters a teacher working with a student. The teacher tells the robot to ``start at the front of the classroom'', expecting it to go to the front of the classroom and begin cleaning that area. This response requires that the robot is able to \emph{determine the spatial region in the environment that satisfies this concept}.

The ability to understand and reason about \textit{spatial regions} is essential for cognitive systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current cognitive systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If cognitive systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Creating systems able to integrate such a range of information is a key challenge in the cognitive systems paradigm \cite{langley:inpress}.

@@

We use \textit{anchor points}~\cite{Klenk/etal2005} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate. Anchor points are linked to particular CDSRs using a \fw{boundarySegment} ternary relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary segments.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy~\cite{Kuipers:2000}. This paper follows in this tradition, adding CDSRs to his qualitative topological representations. Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms. This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}. The need for an autonomous system to ground references to human-generated descriptions of space has been recognized in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.

There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2005}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.

@@

In this paper we presented an integrated cognitive system capable of representing and reasoning about context-dependent spatial regions. The system identifies CDSRs in previously unseen environments through analogy with a single example. This is a difficult cognitive systems task requiring integration of semantic and geometric knowledge to identify regions as small as 8\% of the room. Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and a lack of transfer post-processing (e.g. comparing the QSRs present in both base and transferred regions) to improve results. In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them. Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to transfer CDSRs that overlap with user-defined regions for 6 out of 7 region types.

@@

\bibliography{aaai12-cdsr}
\bibliographystyle{aaai}
\end{document}

>>>>

\title{Using Anchor Points to Define and Transfer Spatial Regions Based on Context}

\author{%
Matthew Klenk \\ Palo Alto Research Center \\  Palo Alto, CA \\ matthew.klenk@parc.com 
\And % 
Nick Hawes \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ n.a.hawes@cs.bham.ac.uk 
\And % 
Kate Lockwood \\ ITCD Department \\ California State University - Monterey Bay \\ klockwood@csumb.edu  
\AND % 
Graham S. Horn \\ Intelligent Robotics Lab \\ University of Birmingham, UK \\ gsh148@cs.bham.ac.uk 
\And % 
John D. Kelleher \\ Applied Intelligence Research Centre \\ Dublin Institute of Technology \\ john.d.kelleher@dit.ie}

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, AI systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

Consider a janitorial robot cleaning a classroom. While performing this task, it encounters a teacher working with a student. The teacher tells the robot to ``start at the front of the classroom'', expecting it to go to the front of the classroom and begin cleaning that area. This response requires that the robot is able to \emph{determine the spatial region in the environment that satisfies this concept}. Moving from the metric space of the robot's sensors to the symbolic space required for reasoning and language is an important problem for qualitative reasoning.

The ability to understand and reason about \textit{spatial regions} is essential for AI systems performing tasks for humans in everyday environments. Some regions, such as whole rooms and corridors, are defined by clearly perceivable boundaries (e.g. walls and doors). However, many regions to which humans routinely refer are not so easily defined. Consider, for example, the aforementioned region \textit{the front of the classroom}. This region is not perceivable using just the geometry of the environment. Instead, it is defined by the objects present in the room (chairs, a desk, a whiteboard), their role in this context (seats for students to watch a teacher who writes on the whiteboard) and their configuration in space (the seats point toward the whiteboard). We refer to such regions as \textit{context-dependent spatial regions} (CDSRs).

Current AI systems are not capable of representing and reasoning about CDSRs, yet it is an important ability. If AI systems are to collaborate with humans in everyday environments then they must be able to understand and refer to the same spatial regions humans do. Many regions are best defined in a context-dependent manner, for example, a kitchen in a studio apartment, an aisle in a church or store, behind enemy lines in a military engagement, etc. In order to represent and reason about such regions, cognitive systems must integrate different types of information, including geometric, semantic, and functional knowledge. Qualitative representations provide a symbolic abstraction that is a natural method for integrating these different types of knowledge in reasoning tasks.

@@

We use \textit{anchor points}~\cite{Klenk/etal2011} to define the boundaries of CDSRs. Anchor points are symbolic descriptions which link a conceptual entity to a perceived entity. The perceived entities we use are the objects recognised by Dora, and the room itself. The room representation is created by putting a convex hull around the lines in Dora's SLAM map. Anchor points are created from perceived entities using unary functions, e.g. \fw{(XMaxYMostFn Desk1)} represents the point on the \fw{Desk1} with the largest x coordinate taken from the set of points with a y coordinate within 5\% of the maximum y coordinate. Anchor points are linked to particular CDSRs using a \fw{boundarySegment} ternary relation. After we have defined the boundary of the region, we assign it a semantic label using the \fw{regionType} relation. Therefore, each CDSR has one type and a variable number of boundary segments.

@@

Typical approaches to spatial representation for mobile robots tend to focus on localization, and thus mostly represent the world uniformly without subdivision into meaningful (semantic) units \cite{Thrun02a}. When a more structured representation is required, many turn to Kuipers' Spatial Semantic Hierarchy (SSH) \cite{Kuipers:2000}. In that work, Kuipers distinguishes between large-scale and small scale space based on the robots sensor horizon. This distinction has proven useful for performing localization and map building \cite{Beeson01042010}. CDSRs differ from this approach by incorporating non-geometric knowledge into the definition of regions. Whilst mobile robots exist which can determine the type of a room from the objects in it ~\cite{Hanheide/etal:2010a,Galindo/etal:2005a}, they only concern themselves with types of whole rooms, and cannot represent regions within rooms. This is also true for those systems which use some elements of QSR~\cite{aydemir2011icra}. The need for an autonomous system to ground references to human-generated descriptions of space has been recognized in domains where a robot must be instructed to perform a particular task, however existing systems are restricted to purely geometrically-defined regions~\cite{Tellex:2011,Dzifcak/etal:2009,brenneretal07ijcai}.

There is mounting evidence that analogy, operating over structured qualitative representations, can be used to simulate a number of spatial reasoning tasks. Forbus \textit{et al.} showed that analogy between course of action diagrams could be used to identify potential ambush locations in new situations by focusing on only the relevant aspects of sketched battle plans \cite{Forbus/etal2003}. A core contribution of their work was the definition of a \textit{shared similarity constraint} between a spatial reasoning system and its user; where users and spatial reasoning systems agree on the similarities between situations. This has close parallels to what we are trying to accomplish, where a cognitive system is able to reason about context-dependent spatial regions by identifying the same salient features as its human user. The anchor points in our work were originally used in teaching a system how to solve problems from the Bennett Mechanical Comprehension Test that require spatial and conceptual reasoning. For example, identifying which wheelbarrow will be more difficult to lift based on the relative locations of its loads as depicted in a sketch \cite{Klenk/etal2011}. In that work, the anchor points defined the endpoints of lines. We go beyond that result to use anchor points to specify 2D regions.

@@

In this paper, we illustrate how qualitative representations enable the integration of geometric and semantic knowledge required to reason about CDSRs. The system identifies CDSRs in previously unseen environments through analogy with a single example. Our system demonstrates a successful integration of a range of technologies including vision, SLAM, qualitative spatial reasoning and analogy to achieve this task. In order to make this rich collection of components work together, our work takes a number of short-cuts that we plan to address with future work. These include a reliance on the initial orientation of a room in a global coordinate frame, the lack of a mechanism to retrieve relevant rooms from memory (e.g. MAC/FAC \cite{forbus/etal1995}), and manually-encoding the CDSRs instead of extracting them from sketched input. In addition, we must complement our system development work with more comprehensive human studies assessing how people define and use these regions as well as how well anchor points capture them. Despite the preliminary nature of this work, our evaluation demonstrates that the system is able to transfer CDSRs that overlap with user-defined regions for 6 out of 7 region types.

@@

\bibliography{QR12-cdsr}
\bibliographystyle{aaai}
\end{document}

### 2012-04-13 code/scenarios/cdsr/trunk/papers/QR2012/qsr-extraction.tex

These spatial relationships between adjacent objects provide the structure necessary for analogical processing. Generating the relationships in this way (as opposed to, for example, simple coordinate-based thresholding) has the advantage that the presence and absence of relationships is represented on a continuous scale. This provides our representations with the flexibility necessary to manage the variation in perceptual information (i.e. the position of walls and objects) inevitable in human environments and robot perception.

>>>>

These spatial relationships between adjacent objects provide the structure necessary for analogical processing. Generating the relationships in this way (as opposed to, for example, simple coordinate-based thresholding) has the advantage that the presence and absence of relationships is represented on a continuous scale. While similar to the relationships used in fuzzy qualitative trigonometry \cite{Liu200971}, our approach includes distance as well as angle into the weight computation. Weighted relationships provide our representations with the flexibility necessary to manage the variation in perceptual information (i.e. the position of walls and objects) inevitable in human environments and robot perception.
